---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# queryparser

<!-- badges: start -->
<!-- badges: end -->

**queryparser** parses SQL queries into lists of R expressions.

## Installation

Install the released version of **queryparser** from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("queryparser")
```

Or install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ianmcook/queryparser")
```

## Usage

Call the function `parse_query()`, passing a `SELECT` statement enclosed in quotes as the first argument:

```{r}
library(queryparser)

parse_query("SELECT DISTINCT carrier FROM flights WHERE dest = 'HNL'")
```

Queries can include the clauses `SELECT`, `FROM`, `WHERE`, `GROUP BY`, `HAVING`, `ORDER BY`, and `LIMIT`:

```{r}
parse_query(
" SELECT origin, dest,
    COUNT(flight) AS num_flts,
    round(AVG(distance)) AS dist,
    round(AVG(arr_delay)) AS avg_delay
  FROM flights
  WHERE distance BETWEEN 200 AND 300
    AND air_time IS NOT NULL
  GROUP BY origin, dest
  HAVING num_flts > 5000
  ORDER BY num_flts DESC, avg_delay DESC
  LIMIT 100;"
)
```

Set the argument `tidyverse` to `TRUE` to use functions from [tidyverse](https://www.tidyverse.org) packages including [dplyr](https://dplyr.tidyverse.org), [stringr](https://stringr.tidyverse.org), and [lubridate](https://lubridate.tidyverse.org) in the R expressions:

```{r}
parse_query("SELECT COUNT(*) FROM t WHERE x BETWEEN y AND z", tidyverse = TRUE)
```

**queryparser** will translate only explicitly allowed functions and operators, preventing injection of malicious code:

```{r error=TRUE}
parse_query("SELECT x FROM y WHERE system('rm -rf /')")
```

## Current Limitations

**queryparser** does not currently support:

- Joins
- Subqueries
- `OVER` expressions
- `CASE` expressions
- Some SQL functions and operators

The error messages that occur when attempting to parse invalid or unrecognized SQL are often non-informative.

Expressions that include logical operators (such as `IS NULL`) might generate R expressions that do not follow the same order of operations that a SQL engine would follow. When using expressions (not simple column references or literals) as operands to logical operators, enclose these expressions in parentheses to avoid ambiguity.

## Non-Goals

**queryparser** does not aspire to:

- Translate other SQL statements (such as `INSERT` or `UPDATE`)
- Customize translations for different SQL dialects
- Fully validate the syntax of the `SELECT` statements passed to it
